---
title: "p8105_hw6_xl3495"
author: "Xueting Li"
date: "2024-12-01"
output: github_document
---

```{r}
library(tidyverse)
library(modelr)
library(broom)
library(purrr)
set.seed(123)
```



# Problem 1

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```
```{r}
n_bootstrap = 5000

boot_models = weather_df |>
  select(tmax, tmin) |>
  bootstrap(n = n_bootstrap) |>
    mutate(
    model = map(strap, ~ lm(tmax ~ tmin, data = .))
  )

boot_r_squared = boot_models |>
  mutate(glance = map(model, glance)) |>
  unnest(glance) |>
  select(r.squared) |>
   mutate(metric = "R-squared")


boot_coefficients = boot_models |>
  mutate(
    coefficients = map(model, coefficients),
    product = map_dbl(coefficients, ~ .[1] * .[2]),
    log_beta = log(product)
  ) |>
  select(log_beta) |>
  mutate(metric = "log(β0 * β1)")
```

```{r}
combined_data = bind_rows(
  boot_r_squared |> rename(value = r.squared),
  boot_coefficients |> rename(value = log_beta)
)

ggplot(combined_data, aes(x = value, fill = metric)) +
  geom_histogram(bins = 30, alpha = 0.6) +
  facet_wrap(~ metric, scales = "free_x") +  # Facet by 'metric' variable
  theme_minimal() +
  scale_fill_manual(values = c("R-squared" = "lightblue", "log(β0 * β1)" = "lightgreen")) +
  labs(
    title = "Distribution of R-squared and log(β0 * β1) from Bootstrap Samples",
    x = "Value", y = "Frequency"
  )
```
From the plot, the distributions of the two estimates follow an approximate normal distribution. For $log(\beta_0 * \beta_1)$, it frequently distributes between (2, 2.025). For, $r^2$,it frequently distributes between (0.905, 0.915).

```{r}
r_squared_ci = quantile(boot_r_squared$r.squared, probs = c(0.025, 0.975))

log_beta_ci = quantile(boot_coefficients$log_beta, probs = c(0.025, 0.975))

ci_df = data.frame(
  Metric = c("R-squared", "log(β0 * β1)"),
  Lower_bound = c(r_squared_ci[1], log_beta_ci[1]),
  Upper_bound = c(r_squared_ci[2], log_beta_ci[2])
)

print(ci_df)
```


# Problem 2

```{r}
homicide_data = read_csv("datasets/homicide-data.csv") |>
  mutate(
    city_state = paste(city, state, sep = ", "),
    homicide_solved = if_else(disposition == "Closed by arrest", 1, 0),
    victim_age = as.numeric(victim_age)
    ) |>
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
    victim_race %in% c("White", "Black"),
    !is.na(victim_age)
    )
```
```{r}
baltimore_data = homicide_data |>
  filter(city_state == "Baltimore, MD")

log_reg_model = glm(homicide_solved ~ victim_age + victim_sex + victim_race, 
                     data = baltimore_data, 
                     family = binomial)

model_summary = tidy(log_reg_model, conf.int = TRUE)

# Extract the estimate and confidence interval for victim_sex variable
sex_results = model_summary |>
  filter(term == "victim_sexMale")

odds_ratio = exp(sex_results$estimate)
lower_ci = exp(sex_results$conf.low)
upper_ci = exp(sex_results$conf.high)

result_table = tibble(
  Variable = "Male vs Female Victims",
  Odds_Ratio = odds_ratio,
  Lower_CI = lower_ci,
  Upper_CI = upper_ci
)

print(result_table)
```


```{r}
city_log_reg = homicide_data |>
  filter(
    victim_sex %in% c("Male", "Female"), victim_race %in% c("White", "Black")
    ) |>
  group_by(city) |>
  nest() |>
  mutate(
    model = map(data, ~ glm(homicide_solved ~ victim_age + victim_sex + victim_race, 
                            data = ., family = binomial)),
    tidy_results = map(model, tidy, conf.int = TRUE)
  )


city_or_results = city_log_reg |>
  unnest(tidy_results) |>
  filter(term == "victim_sexMale") |>
  mutate(
    odds_ratio = exp(estimate),
    lower_ci = exp(conf.low),
    upper_ci = exp(conf.high)
  ) |>
  select(city, odds_ratio, lower_ci, upper_ci) |>
  arrange(desc(odds_ratio))

head(city_or_results)
```


```{r}
ggplot(city_or_results, aes(x = reorder(city, odds_ratio), 
                            y = odds_ratio, 
                            ymin = lower_ci, 
                            ymax = upper_ci)) +
  geom_point() + 
  geom_errorbar(width = 0.3) +
  coord_flip() +
  labs(
    title = "Adjusted Odds Ratio for Solving Homicides: Male vs Female Victims by City",
    x = "City",
    y = "Adjusted Odds Ratio (Male vs Female)"
  ) +
  theme_minimal()
```

According to the plot,  if the point is above 1, it suggests that male victims have a higher likelihood of having their homicide solved compared to female victims in that city, and if the CI includes 1, the effect is not statistically significant, meaning we cannot conclude whether there's a difference in the likelihood of solving homicides for male vs. female victims in that city. For example, male victims in Albuquerque have a higher likelihood of having their homicide solved compared to female victims. There are still a number of cities we cannot conclude whether there's a difference in the likelihood of solving homicides for male vs. female victims.

# Problem 3




